Homework 2
====================

Preprocessing
-------------

### Load Data
Includes helper function to calculate Misclassification error rate.
```{r}
setwd("~/Users/SMW/Dropbox/Coursera/DA_012213/Homework/HW2/code")
source('mcErrorRate.R')
##download.file("https://spark-public.s3.amazonaws.com/dataanalysis/samsungData.rda", "../data/samsungData.rda")
```

### Examine raw data
```{r}
names(samsungData)
table(is.na(samsungData)) ## no NAs
table(samsungData$subject)
```
There are no NAs found.
21 subjects total, with numeric identifiers.

### Data Munging
Make variable names syntactically correct by removing parentheses, commas, etc.
```{r}
nameVec <- make.names(names(samsungData),unique=TRUE)
names(samsungData) <- nameVec
names(samsungData)
```


### make Training, and Test sets
Training and test sets include the require subjects as noted.
training set size has 17 subjects.
Test set contains the required 4 subjects only.
```{r}
test1 <- c(27,28,29,30)
train1 <-c(1,3,5,6)
rest <- unique(samsungData$subject)
rest <- setdiff(setdiff(rest,test1),train1)
set.seed(35252)
test1 <- c(test1, sample(rest,3, replace=FALSE))
train1 <- setdiff(c(train1,rest),test1)
trainSet <- samsungData[!(samsungData$subject %in% train1),]
testSet <- samsungData[samsungData$subject %in% test1,]
table(trainSet$subject)
table(testSet$subject)
```

Pick Function
-------------
### Classification tree.
First, create a classification tree using all variables as covariates.
```{r}
library(tree)
train.tree <- tree(factor(activity)~.,data=trainSet)
par(mfrow=c(1,1))
plot(train.tree)
text(train.tree, cex=0.5)
summary(train.tree)
```
The resulting tree seems to indicate that energy and gravitational acceleration are good predictors. Why?
1. When stationary, low energy. when moving, high energy.
2. Gravitational acceleration is constant. When moving, the phone orientation changes, so possibly the mean acceleration for a particular axis would be lower than when stationary.
3. Jerk should be more significant when moving up and down.


#### Cross-validate tree
```{r}
par(mfrow=c(2,1))
plot(cv.tree(train.tree,FUN=prune.tree,method="misclass"))
plot(cv.tree(train.tree))
cv.tree(train.tree)
```
Resulting tree is OK but not great. It seems to indicate an interest it gravity!

#### Prune tree
Try 6 nodes first, since 6 outcomes


```{r}
tree.predict <- predict(train.tree, trainSet[,-563])
head(tree.predict)
table(trainSet$activity, predict(train.tree, type="class"))
```


### random Forest
```{r}
library(randomForest)
set.seed(37043)
train.rf <- randomForest(factor(activity) ~., data=trainSet, prox=TRUE)
print(train.rf)
importance(train.rf)
par(mfrow=c(1,1))
varImpPlot(train.rf, main="Importance Plot", cex=0.7)
getTree(train.rf,k=1)
table(validateSet$activity,predict(train.rf,validateSet[,-563]))
```

Cross-validate
--------------

Apply to test set
-----------------
```{r}
table(testSet$activity,predict(train.rf,testSet[,-563]))
```

### Refine if validation